Fish & Benthic Community data: Antigua, Barbuda, & Bonaire - Aug. 2017

### 2018-11-14 r/SEM.Rmd
11/13

```{r,message=FALSE,warning=FALSE, echo=FALSE}
# set up
library(rmarkdown)
library(dplyr)
library(tidyr)
library(lavaan)
library(semPlot)
```

```{r,message=FALSE,warning=FALSE, echo=FALSE}
setwd("~/github/Scarid-Behavior/data")
site.sum <- read.csv("site.sum.csv", stringsAsFactors=FALSE) %>% select(-X)
data <- read.csv("data2.csv", stringsAsFactors=FALSE)
# add aglal indeces
data <- data %>% mutate(index_MA = cover_MA*canopy_MA) %>% mutate(index_TA = cover_TA*canopy_TA)
# recode phase and species with dummy variables
data <- data %>% mutate(spec.dum= 
                          ifelse(Species.Code=="stop", 1, 
                          (ifelse(Species.Code=="qup", 2,
                          (ifelse(Species.Code=="rbp",3,NA)))))) %>%
  mutate(phase.dum=ifelse(Phase=="i",1,2))

@@

```{r}
# define model pathways
model <- '
    FR ~ Length.cm + phase.dum + spec.dum+ int.scar + RISK + NUTRITION
    int.scar ~ scar.BM + cover_MA + cover_TA
    rugosity ~ cover_LC
    cover_LC ~ cover_MA + cover_TA
    cover_TA ~ scar.BM
    cover_MA ~ scar.BM

    RISK =~ scar.BM + rugosity
    NUTRITION =~ cover_MA + cover_TA
'

fit <- sem(model, data=data)
varTable(fit)
summary(fit, standardized=T, fit.measures=T,rsq=T)
semPaths(fit)
```

>>>>

```{r,message=FALSE,warning=FALSE, echo=FALSE}
# set up
library(tidyr)
library(lavaan)
library(semPlot)
```

```{r,message=FALSE,warning=FALSE, echo=FALSE}
sum_id <- read.csv(here("data","sum_id.csv"), stringsAsFactors=FALSE) %>% select(-X)
sum_site <- read.csv(here("data","sum_site.csv"), stringsAsFactors=FALSE) %>% select(-X)

# recode phase and species with dummy variables
# data <- data %>% mutate(spec.dum= 
#                           ifelse(Species.Code=="stop", 1, 
#                           (ifelse(Species.Code=="qup", 2,
#                           (ifelse(Species.Code=="rbp",3,NA)))))) %>%
#   mutate(phase.dum=ifelse(Phase=="i",1,2))

@@

```{r}
# define model pathways
qup_df <- sum_id %>% 
  select(scar_bm, ma_cover, ta_cover, fr, g_frac, rugosity, int_scar, length_cm, lc_cover) %>% 
  na.omit()

sem <- '
  # latent variable indicators  
      risk =~ scar_bm + rugosity
      nutrition =~ ma_cover + ma_canopy + ta_cover + ta_canopy
  # regressions
      fr ~ length_cm + risk + nutrition
      rugosity ~ lc_cover
      lc_cover ~ ma_cover + ta_cover
      ta_cover ~ scar_bm
      ma_cover ~ scar_bm
'

sem1 <- '
  # latent variable indicators
      risk =~ scar_bm      
      nutrition =~ ma_cover + ta_cover
  # regressions
      fr ~ length_cm + nutrition + risk
      lc_cover ~ ma_cover + ta_cover
'

fit <- sem(sem1, data=qup_df)
varTable(fit)
summary(fit, standardized=T, fit.measures=T,rsq=T)
semPaths(fit)
fitMeasures(fit, c("cfi","rmsea","srmr"))
```

### 2018-11-14 r/Sites.Rmd
11/13

Notes 11/11:
* figure out dcast issues
* replacing NA/0s within pipe

@@

## Total fish populations, by species and site
```{r,message=F,warning=F}
# Because not all species were present at each site/transect, I need to create template to add 0 values for biomass and density in transects where they are absent.
exp_tran <- fish %>% 
  group_by(site,transect) %>% 
  summarize() # all possible site/transect combinations
exp_sp <- fish %>% expand(nesting(species,family),site) # all possible species/site combinations
template_sp <- full_join(exp_sp,exp_tran,by="site") # species/site/transect combinations

@@

# Scarids by phase (should this be by species, too?)
exp_ph <- fish %>% expand(phase,site)
template_ph <- full_join(exp_ph,exp_tran,by="site") # all site/transect/phase combinations

@@

```{r, echo=F,message=F,warning=F}
# Because not all categories were present within each meter, I need to create template to add 0 values in meters where they were ab_fishsent.
exp_meter <- ab_benthic %>% 
  group_by(site,transect,meter) %>% 
  summarize()
exp_cat <- ab_benthic %>% expand(nesting(code),site)
template_bent <- full_join(exp_meter,exp_cat,by="site")

@@

bent_site <- bent_site %>% 
      spread_n(code, c(cover, cover_se)) %>%
  left_join(can_site %>% 
            spread_n(code, c(canopy,canopy_se))
  ) %>% 
  janitor::clean_names(case="old_janitor")

@@

bent_site <- bent_site %>% left_join(rug_site,by="site") %>% bind_rows(bon_benthic)
```

>>>>

Wrangling of fish, benthic, and rugosity data from Antigua, Barbuda, and Bonaire to put in usable site-level format.

@@

## Total fish populations, by species and site
```{r,message=F,warning=F}
# Because not all species were present at each site/transect, I need to create template to add 0 values for biomass and density in transects where they are absent.
exp_tran <- fish %>% 
  group_by(site,transect) %>% 
  summarize() # all possible site/transect combinations
exp_sp <- fish %>% tidyr::expand(nesting(species,family),site) # all possible species/site combinations
template_sp <- full_join(exp_sp,exp_tran,by="site") # species/site/transect combinations

@@

# Scarids by phase (should this be by species, too?)
exp_ph <- fish %>% tidyr::expand(phase,site)
template_ph <- full_join(exp_ph,exp_tran,by="site") # all site/transect/phase combinations

@@

```{r, echo=F,message=F,warning=F}
# Because not all categories were present within each meter, I need to create template to add 0 values in meters where they were ab_fishsent.
exp_meter <- ab_benthic %>% 
  group_by(site,transect,meter) %>% 
  summarize()
exp_cat <- ab_benthic %>% tidyr::expand(nesting(code),site)
template_bent <- full_join(exp_meter,exp_cat,by="site")

@@

bent_site <- bent_site %>% 
      spread_n(code, c(cover, cover_se)) %>%
  left_join(can_site %>% 
            spread_n(code, c(canopy,canopy_se))
  ) %>% 
  janitor::clean_names(case="old_janitor") %>%
  select(cca_cover, cca_cover_se, lc_cover, lc_cover_se, ma_cover, ma_cover_se, ma_canopy, ma_canopy_se, ta_cover, ta_cover_se, ta_canopy, ta_canopy_se)

@@

bent_site <- bent_site %>% 
  left_join(rug_site,by="site") %>% 
  bind_rows(bon_benthic) %>%
  mutate(ma_ind = ma_canopy*ma_cover,
         ta_ind = ta_canopy*ta_cover)
```
